---
title: "Parametric Bootstrapping"
author: "CJ Carani, Dylan Hunt and Matt Kaznikov"
date: "02-21-2024"
output: pdf_document
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE, fig.height = 3, fig.width = 6)
library(tidyverse)
library(GGally)
library(knitr)
library(gridExtra)
library(MASS)
library(gridExtra)  
library(mnormt) 
library(lme4) 
library(lmerTest)
library(knitr) 
library(kableExtra)
library(tidyverse)
library(Matrix)
```

```{r}
Airbnb <- read_csv("https://raw.githubusercontent.com/proback/BeyondMLR/master/data/airbnb.csv")
```

<<<<<<< HEAD
=======
Use the Airbnb to compare models
maybe llsr vs lme?
```{r}
#llsr model
m1 <- lm(data = Airbnb, price ~ BikeScore + WalkScore + TransitScore)

#lme model
m2 <- lmer(data = Airbnb, price ~ BikeScore + WalkScore + TransitScore + (1|neighborhood))
```

Data for a model without many observations (Elephants)

```{r}
elephants <- read_csv("https://raw.githubusercontent.com/proback/BeyondMLR/master/data/elephant.csv")
```

```{r}
E_M1 <- lm(MATINGS ~ AGE, data = elephants)
summary(E_M1)

elephants <- elephants %>% mutate(age2 = AGE*AGE)
E_M2 <- lm(MATINGS ~ AGE + age2, data = elephants)
summary(E_M2)
```
#Bootstrap for Model 1 Estimates
```{r}
# record estimates from the model
b0 <- E_M1$coefficients[1]
b1 <- E_M1$coefficients[2]
sigma <- summary(E_M1)$sigma
nreps <- 10000
bootstrap_b0 <- c(rep(NA, nreps))
bootstrap_b1 <- c(rep(NA, nreps))
bootstrap_sigma <- c(rep(NA, nreps))
E_Data <- elephants
# simulate data and fit model
for (i in 1:nreps){
   # simulate new data
E_Data <- E_Data %>% mutate(SimMatings = b0 + b1*AGE + rnorm(n= nrow(elephants), mean=0, sd=sigma))
Mb <- lm(data=E_Data, SimMatings~AGE)   #fit model to simulated data
bootstrap_b0[i] <- Mb$coefficients[1]        #record b0 from model on simulated data
bootstrap_b1[i] <- Mb$coefficients[2]        #record b1 from model on simulated data
bootstrap_sigma[i] <- summary(Mb)$sigma      #record sigma from model on simulated data
}
Bootstrap_Results <- data.frame(bootstrap_b0, bootstrap_b1, bootstrap_sigma)
```

```{r}
p1 <- ggplot(data=Bootstrap_Results, aes(x=bootstrap_b1)) + geom_histogram() + geom_vline(xintercept=quantile(bootstrap_b1, c(.025, .975)), color="red")
p1
```

```{r}
quantile(bootstrap_b1, c(.025, .975))
```

```{r}
# record estimates from the model
b0 <- E_M2$coefficients[1]
b1 <- E_M2$coefficients[2]
b2 <- E_M2$coefficients[3]
sigma <- summary(E_M2)$sigma
nreps <- 10000
bootstrap_b0 <- c(rep(NA, nreps))
bootstrap_b1 <- c(rep(NA, nreps))
bootstrap_b2 <- c(rep(NA, nreps))
bootstrap_sigma <- c(rep(NA, nreps))
E_Data <- elephants
# simulate data and fit model
for (i in 1:nreps){
   # simulate new data
E_Data <- E_Data %>% mutate(SimMatings = b0 + b1*AGE + b2*age2 + rnorm(n= nrow(elephants), mean=0, sd=sigma))
Mb <- lm(data=E_Data, SimMatings~AGE)        #fit model to simulated data
bootstrap_b0[i] <- Mb$coefficients[1]        #record b0 from model on simulated data
bootstrap_b1[i] <- Mb$coefficients[2]        #record b1 from model on simulated data
bootstrap_b2[i] <- Mb$coefficients[3]        #record b2 from model on simulated data
bootstrap_sigma[i] <- summary(Mb)$sigma      #record sigma from model on simulated data
}
Bootstrap_Results <- data.frame(bootstrap_b0, bootstrap_b1, bootstrap_b2, bootstrap_sigma)
```

```{r}
quantile(bootstrap_b2, c(.025, .975))
```

```{r}
anova(E_M1, E_M2)
```


```{r}
set.seed(02272022)
# record estimates from reduced model
b0 <- E_M1$coefficients[1]
b1 <- E_M1$coefficients[2]
sigma <- summary(E_M1)$sigma
nreps <- 10000
bootstrap_F <- c(rep(NA, nreps))
B_Data <- elephants
# simulate data and fit model
for (i in 1:nreps){
   # simulate new data from reduced model
B_Data <- B_Data %>% mutate(SimMatings = b0 + b1*AGE + rnorm(n= nrow(elephants), mean=0, sd=sigma))
Mb_Red <- lm(data=B_Data, SimMatings~AGE)   #fit reduced model to simulated data
Mb_Full <- lm(data=B_Data, SimMatings ~ AGE + age2)   #fit full model to simulated data
bootstrap_F[i] <- anova(Mb_Full, Mb_Red)$F[2]
}
Bootstrap_Results <- data.frame(bootstrap_F)
```

```{r}
p1 <- ggplot(data=Bootstrap_Results, aes(x=bootstrap_F)) + geom_histogram() + geom_vline(xintercept = 0.4187, color="red", linetype="dotted", size=2) 
p1
```



```{r}
## Multilevel Generalized Linear Model for Bootstrapping

# Reading in Basketball Refs Data
refdata <- read.csv("https://raw.githubusercontent.com/proback/BeyondMLR/master/data/basketball0910.csv")
# Model 1 with random slope, but no random Effect
glm1 <- glmer(foul.home ~ foul.diff + (1|game), family = binomial(link="logit"), data = refdata)
# Model 2 with a random slope and a random effect for foul.diff
glm2 <- glmer(foul.home ~ foul.diff + (foul.diff|game), family = binomial(link="logit"), data = refdata)

# Compare Model's
anova(glm1,glm2)
```

```{r}
# Bootstrapping
set.seed(02272022)
nreps <- 100
ChiSq <- rep(NA, nreps)
for(i in 1:nreps){
SimData <- drop(simulate(glm1))  # this command simulates data directly from a model
glm1B <-refit(glm1, newresp=SimData)  # refits glm1 to simulated data
glm2B <-refit(glm2, newresp=SimData)  # refits glm2 to simulated data
ChiSq[i] <- anova(glm1B,glm2B)$Chisq[2]
}
ChiSq<- write.csv(ChiSq, file="ChiSq.csv")
```

```{r}
p <- ggplot(data=data.frame(ChiSq), aes(x=ChiSq)) + geom_histogram() + 
  geom_vline(xintercept = 5.4682, color="red", linetype="dotted", size=2) 
p
```

```{r}
mean(ChiSq>5.4682)
```

Level 1 Variables:
  Observational Unit: Fouls
  Variables:
    foul.num, foul.home, foul.vis, foul.diff, score.diff, lead.vis, lead.home, previous.foul.home,          previous.foul.vis, foul.type, personal, offensive, time
Level 2 Variables:
  Observational Unit: Game
  Variables:
    Game, Date, Visitor, Home Team
    

=======
# Background

# Data

# Methods

# Results

# Conclusions
